Cloudbreak: A MapReduce Algorithm for Detecting Genomic Structural Variation

### 2013-02-15 manuscript/manuscript.tex
everything up-to-date with new simulation

% latex table generated in R 2.15.2 by xtable 1.7-0 package
% Tue Jan  1 22:36:02 2013
\begin{table}[t]
\begin{center}
\begin{tabular}{rrr|rrrrr}
  \hline
 & Prec. & Recall & 40-100bp & 101-250bp & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
Total Number & & & 9,980 & 242 & 218 & 110 & 375 \\
  \hline
Cloudbreak & 0.100 & \textbf{0.090} & \textbf{436} (174) & 128 (\textbf{9}) & 158 (\textbf{8}) &  70 (\textbf{4}) & 186 (12) \\ 
Breakdancer & 0.120 & 0.086 & 260 (43) & \textbf{132} (8) & \textbf{167} (2) & \textbf{92} (1) & \textbf{288} (\textbf{14}) \\
  GASVPro & \textbf{0.127} & 0.048 & 102 (11) &  37 (2) & 77 (0) &  26 (0) & 93 (0) \\ 
  DELLY & 0.081 & 0.070 & 144 (10) & 125 (8) & 158 (1) &  83 (1) & 256 (3) \\ 
  Pindel & 0.013 & 0.061 & 414 (\textbf{297}) & 40 (0) & 90 (0) &  37 (0) & 88 (0) \\ 
   \hline
\end{tabular}
\end{center}
\caption{The precision and recall with respect to the gold standard set of each tool on the NA18507 data, as well as the number of deletions in each size class found. The same cutoffs were used as for the simulated data as in Table \ref{chr2preds}, adjusted for the difference in coverage from 30X to 37X.}
\label{NA18507preds}
\end{table}

@@

We analyzed the predictions made by Cloudbreak to determine whether any particular class of deletion event was contributing to its performance. We found that Cloudbreak shows particularly strong performance in detecting deletions that overlap RepeatMasker-annotated elements (Table \ref{repmaskpreds}) in both the simulated and NA18507 data sets, especially compared to the RP-based methods. In the simulated data set, Cloudbreak correctly identifies 9\% more repeat-overlapping deletions than the nearest competitor. Cloudbreak identifies more deletions in repeats than any other methods, and finds almost four times more exclusive hits than the other RP-based methods. Since deletions and other SVs are enriched in repetitive regions of the genome in both of our data sets, this contributes greatly to Cloudbreak's performance, and demonstrates the value in considering as many possible mappings of ambiguously mapping reads as possible.

\begin{table}[b]
\begin{center}
\begin{tabular}{rrr|rr}
 & \multicolumn{2}{c}{Simulated Data} & \multicolumn{2}{c}{NA18507} \\
\hline
 &  Non-repeat & Repeat  &  Non-repeat & Repeat \\ 
 Total Number & 122 & 317 & 604 & 10,321 \\ 
  \hline
Cloudbreak & \textbf{29} (4) & \textbf{155} (\textbf{19}) & \textbf{204} (\textbf{38}) & \textbf{774} (169) \\ 
  Breakdancer &  \textbf{29} (\textbf{6}) & 142 (10) & 186 (25) & 753 (43) \\ 
  GASVPro & 15 (2) & 72 (2) & 71 (5) & 264 (8) \\ 
  DELLY & 21 (2) & 116 (2) & 147 (12) & 619 (11) \\ 
  Pindel & 8 (4) & 62 (12) & 98 (26) & 571 (\textbf{271}) \\ 
   \hline
\end{tabular}
\end{center}
\caption{Detected deletions on the simulated and NA18507 data sets identified by each tool, broken down by whether the deletion overlaps with a RepeatMasker-annotated element.}
\label{repmaskpreds}
\end{table}

>>>>

% latex table generated in R 2.15.2 by xtable 1.7-0 package
% Tue Jan  1 22:36:02 2013
\begin{table}[t]
\begin{center}
\begin{tabular}{rrr|rrrrr}
  \hline
 & Prec. & Recall & 40-100bp & 101-250bp & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
Total Number & & & 7,466 & 235 & 218 & 110 & 375 \\
  \hline
Cloudbreak & 0.098 & \textbf{0.115} & \textbf{423} (163) & 128 (\textbf{9}) & 158 (\textbf{8}) &  70 (\textbf{4}) & 186 (12) \\ 
Breakdancer & 0.097 & 0.111 & 258 (42) & \textbf{132} (8) & \textbf{167} (2) & \textbf{92} (1) & \textbf{288} (\textbf{14}) \\
  GASVPro & \textbf{0.132} & 0.040 & 102 (11) &  37 (2) & 77 (0) &  26 (0) & 93 (0) \\ 
  DELLY & 0.081 & 0.091 & 143 (10) & 125 (8) & 158 (1) &  83 (1) & 256 (3) \\ 
  Pindel & 0.013 & 0.079 & 406 (\textbf{290}) & 40 (0) & 90 (0) &  37 (0) & 88 (0) \\ 
   \hline
\end{tabular}
\end{center}
\caption{The precision and recall with respect to the gold standard set of each tool on the NA18507 data, as well as the number of deletions in each size class found. The same cutoffs were used as for the simulated data as in Table \ref{chr2preds}, adjusted for the difference in coverage from 30X to 37X.}
\label{NA18507preds}
\end{table}

@@

We analyzed the predictions made by Cloudbreak to determine whether any particular class of deletion event was contributing to its performance. We found that Cloudbreak shows particularly strong performance in detecting deletions that overlap RepeatMasker-annotated elements (Table \ref{repmaskpreds}) in both the simulated and NA18507 data sets, especially compared to the RP-based methods. In the simulated data set, Cloudbreak correctly identifies 9\% more repeat-overlapping deletions than the nearest competitor. Cloudbreak identifies more deletions in repeats than any other method, and in the NA18507 data set finds 3.75 times more exclusive hits than the other RP-based methods. Since deletions and other SVs are enriched in repetitive regions of the genome in both of our data sets, this contributes greatly to Cloudbreak's performance, and demonstrates the value in considering as many possible mappings of ambiguously mapping reads as possible.

\begin{table}[b]
\begin{center}
\begin{tabular}{rrr|rr}
 & \multicolumn{2}{c}{Simulated Data} & \multicolumn{2}{c}{NA18507} \\
\hline
 &  Non-repeat & Repeat  &  Non-repeat & Repeat \\ 
 Total Number & 122 & 317 & 604 & 10,321 \\ 
  \hline
Cloudbreak & \textbf{29} (4) & \textbf{155} (\textbf{19}) & \textbf{204} (\textbf{38}) & \textbf{761} (158) \\ 
  Breakdancer &  \textbf{29} (\textbf{6}) & 142 (10) & 186 (25) & 751 (42) \\ 
  GASVPro & 15 (2) & 72 (2) & 71 (5) & 264 (8) \\ 
  DELLY & 21 (2) & 116 (2) & 147 (12) & 618 (11) \\ 
  Pindel & 8 (4) & 62 (12) & 98 (26) & 563 (\textbf{264}) \\ 
   \hline
\end{tabular}
\end{center}
\caption{Detected deletions on the simulated and NA18507 data sets identified by each tool, broken down by whether the deletion overlaps with a RepeatMasker-annotated element.}
\label{repmaskpreds}
\end{table}
